---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MMRFBiolinks 
## An R package that extends TCGABiolink package for integrative analysis with MMRF-COMMPASS data

<!-- badges: start -->
<!-- badges: end -->

MMRFBiolinks extends TCGABiolink package for searching, downloading and analyzing MMRF-COMMPASS data available at the NCI’s Genomic Data Commons (GDC) Data Portal.

## Installation

Once R (version “4.0”) has been started, you can install the released version of MMRFBiolinks from GitHub with:

``` r
devtools::install_github("marzia.settino/MMRFBiolinks", build_vignettes = TRUE)
library(MMRFBiolinks)
```

## Required libraries


```{r eval=FALSE, message=FALSE, warning=FALSE}
library(TCGAbiolinks)
library(SummarizedExperiment)
library(dplyr)
library(DT)
library(ggplot2)
```


## Vignettes
A list of all currently integrated vignettes can be obtained through:



```{r eval=FALSE, message=FALSE, warning=FALSE}
vignette(package="MMRFBiolinks")
```
The best way to view vignettes is in your web browser:


```{r eval=FALSE, message=FALSE, warning=FALSE}
browseVignettes("MMRFBiolinks")
```

Get the list of the example data sets

```{r eval=FALSE, message=FALSE, warning=FALSE}
data(package = "MMRFBiolinks")
```

## Citation





