---
title: "Downloading and preparing files for analysis"
package: MMRFBiolinks
output:
  BiocStyle::html_document
vignette: |
  %\VignetteIndexEntry{3. Downloading and preparing files for analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
library(TCGAbiolinks)
library(MMRFBiolinks)
library(SummarizedExperiment)
library(dplyr)
library(DT)
```




<div class="panel panel-info">
<div class="panel-heading"><b>MMRFGDC_prepare</b></div>
<div class="panel-body">

When using the function <b>MMRFBiolinks::MMRFGDC_prepare</b> there is an argument called SummarizedExperiment which defines the output type <em> Summarized Experiment</em> (default option) or a data frame.


</div>
</div>

<div class="panel panel-warning">
<div class="panel-heading">Summarized Experiment</div>
<div class="panel-body">


[SummarizedExperiment object](http://www.nature.com/nmeth/journal/v12/n2/fig_tab/nmeth.3252_F2.html) contains one or more assays, each represented by a matrix-like object of numeric or other mode. The rows typically represent genomic ranges of interest and the columns represent samples.

A SummarizedExperiment object has three main matrices that can be accessed using the [SummarizedExperiment package](http://bioconductor.org/packages/SummarizedExperiment/)): 

- Sample matrix information is accessed via `colData(data)`: stores sample information and further indexed clinical data and subtype information;
- Assay matrix information is accessed via `assay(data)`: stores molecular data:
- Feature matrix information (gene information) is accessed via `rowRanges(data)`: stores metadata about the features, including their genomic ranges.


MMRFGDC_prepare transforms the downloaded data into a summarizedExperiment object or a data frame. If
summarizedExperiment is set to TRUE, metadata will add to the object. 

If a summarizedExperiment object was chosen, data can be accessed with three different accessors: assay for
the data information, rowRanges to gets the range of values in each row and colData to get the sample information 
(patient, batch, sample type, etc)
8,9



</div>
</div>


# Downloading and preparing data for analysis

## MMRFGDC\_prepare

<b>MMRFGDC\_prepare</b> allows the user to prepare the gene expression data into an R object  for futher analyses.The useful arguments for preparing data from MMRF-CoMMpass Project data are:


 
 
| Argument 	| Description	|
|-------------------------------	|-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------	|
| query 	| A query for GDCquery function 	|
| save 	| Save result as RData object? 	|
| save.filename 	| Name of the file to be save if empty an automatic will be created 	|
| directory 	| Directory/Folder where the data was downloaded. Default: GDCdata 	|
| summarizedExperiment 	| Create a summarizedExperiment? Default TRUE (if possible) 	|
| remove.files.prepared 	| Remove the files read? Default: FALSE This argument will be considered only if save argument is set to true 	|
| add.gistic2.mut 	| If a list of genes (gene symbol) is given, columns with gistic2 results from GDAC firehose (hg19) and a column indicating if there is or not mutation in that gene (hg38) (TRUE or FALSE - use the MAF file for more information) will be added to the sample matrix in the summarized Experiment object. 	|
| mut.pipeline 	| If add.gistic2.mut is not NULL this field will be taken in consideration. Four separate variant calling pipelines are implemented for GDC data harmonization. Options: muse, varscan2, somaticsniper, MuTect2. For more information: https://gdc-docs.nci.nih.gov/Data/Bioinformatics_Pipelines/DNA_Seq_Variant_Calling_Pipeline/ 	|
| mutant_variant_classification 	| List of mutant_variant_classification that will be consider a sample mutant or not. Default: "Frame_Shift_Del", "Frame_Shift_Ins", "Missense_Mutation", "Nonsense_Mutation", "Splice_Site", "In_Frame_Del", "In_Frame_Ins", "Translation_Start_Site", "Nonstop_Mutation" 	|

Example:

```{r results = 'hide', message=FALSE, warning=FALSE, eval = F}

query.mm<-GDCquery(project = "MMRF-COMMPASS",
                    data.category = "Transcriptome Profiling",
                    data.type = "Gene Expression Quantification",
                    workflow.type="HTSeq - FPKM",
                    barcode = c("MMRF_2473","MMRF_2111",
                                "MMRF_2362","MMRF_1824"))

GDCdownload(query.mm, method = "api", files.per.chunk = 10)
data <- MMRFGDC_prepare(query.mm)
```


```{r message=FALSE, warning=FALSE, include=FALSE}
data <- mm.exp.bar
```

```{r  echo=TRUE, message=FALSE, warning=FALSE}




datatable(as.data.frame(colData(data)), 
              options = list(scrollX = TRUE, keys = TRUE, pageLength = 5), 
              rownames = FALSE)

# Only first 100 to make faster
datatable(assay(data)[1:100,], 
              options = list(scrollX = TRUE, keys = TRUE, pageLength = 5), 
              rownames = TRUE)
rowRanges(data)

```




