---
title: "Introduction"
date: "`r BiocStyle::doc_date()`"
author:
- name: Marzia Settino
  affiliation: Data Analytics Research Center
               Department of Medical and Surgical Sciences
               Magna Graecia University,
               Catanzaro, Italy.
- name: Second Author
  affiliation: Data Analytics Research Center
               Department of Medical and Surgical Sciences
               Magna Graecia University,
               Catanzaro, Italy.
  email: marzia.settino@studenti.unicz.it
package: MMRFBiolinks
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{1. Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MMRFBiolinks)
```
MMRFBiolinks is an extension of [TCGAbiolinks package](https://bioconductor.org/packages/TCGAbiolinks/) 
for searching, downloading and preparing data from the Multiple Myeloma Research Foundation (MMRF) CoMMpass study available at the NCI's Genomic Data Commons (GDC) Data Portal. 



# Citation
If you use MMRFBiolinks, please cite:


<hr>
# Download and Install
<hr>
To install this package, start R (version "4.0.0") and enter: 
```R
devtools::install_github("marziasettino/MMRFBiolinks", build_vignettes = TRUE)
library(MMRFBiolinks)
```

# Question and issues
<hr>
Please use [Github issues](https://github.com/MarzyUnicz/MMRFBiolinks/issues) if you want to file bug reports or feature requests


# Required libraries
<hr>
The examples in this tutorial use the following libraries: 
```{r message=FALSE, warning=FALSE, include=TRUE}
library(TCGAbiolinks)
library(SummarizedExperiment)
library(dplyr)
library(DT)
library(ggplot2)
```

# Session info
<hr>
```{r}
version
packageVersion("MMRFBiolinks")
```