---
title: "Introduction"
date: "`r BiocStyle::doc_date()`"
author:
- name: Marzia Settino
  affiliation: Data Analytics Research Center
               Department of Medical and Surgical Sciences
               Magna Graecia University,
               Catanzaro, Italy.
  email: marzia.settino@studenti.unicz.it
- name: Mario Cannataro
  affiliation: Data Analytics Research Center
               Department of Medical and Surgical Sciences
               Magna Graecia University,
               Catanzaro, Italy.
  email: cannataro@unicz.it
package: MMRFBiolinks
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{1. Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=FALSE}
library(MMRFBiolinks)
```
MMRFBiolinks leverages [TCGAbiolinks package](https://bioconductor.org/packages/TCGAbiolinks/) 
for searching, downloading and preparing data from the Multiple Myeloma Research Foundation (MMRF) CoMMpass study available at the NCI's Genomic Data Commons (GDC) Data Portal. 



# Citation
If you use MMRFBiolinks, please cite:

* Settino M, Cannataro M.
MMRFBiolinks: an R-package for integrating and analyzing MMRF-CoMMpass data

* Colaprico A, Silva TC, Olsen C, Garofano L, Cava C, Garolini D, Sabedot T, Malta TM, Pagnotta SM, Castiglioni I, Ceccarelli M, Bontempi G and Noushmehr H. "TCGAbiolinks: an R/Bioconductor package for integrative analysis of TCGA data." Nucleic acids research (2015): gkv1507.

* Mounir, Mohamed, Lucchetta, Marta, Silva, C T, Olsen, Catharina, Bontempi, Gianluca, Chen, Xi, Noushmehr, Houtan, Colaprico, Antonio, Papaleo, Elena (2019). “New functionalities in the TCGAbiolinks package for the study and integration of cancer data from GDC and GTEx.” PLoS computational biology, 15(3), e1006701.

* Silva TC, Colaprico A, Olsen C et al.TCGA Workflow: Analyze cancer genomics and epigenomics data using Bioconductor packages [version 2; peer review: 1 approved, 2 approved with reservations]. F1000Research 2016, 5:1542
(https://doi.org/10.12688/f1000research.8923.2)

[![doi](https://img.shields.io/badge/doi-10.1093/nar/gkv1507-green.svg?style=flat)](http://dx.doi.org/10.1093/nar/gkv1507) [![citation](https://img.shields.io/badge/cited%20by-191-green.svg?style=flat)](https://scholar.google.com.hk/scholar?oi=bibs&hl=en&cites=6029790855238928406) [![Altmetric](https://img.shields.io/badge/Altmetric-25-green.svg?style=flat)](https://www.altmetric.com/details/4919535)



<hr>
# Download and Install
<hr>
To install this package, start R (version "4.0.0") and enter: 
```R
devtools::install_github("marziasettino/MMRFBiolinks", build_vignettes = TRUE)
library(MMRFBiolinks)
```

# Question and issues
<hr>
Please use [Github issues](https://github.com/marziasettino/MMRFBiolinks/issues) if you want to file bug reports or feature requests


# Required libraries
<hr>
The examples in this tutorial use the following libraries: 
```{r message=FALSE, warning=FALSE, include=TRUE}
library(TCGAbiolinks)
library(SummarizedExperiment)
library(dplyr)
library(DT)
library(ggplot2)
```

# Session info
<hr>
```{r}
version
packageVersion("MMRFBiolinks")
```